\lhead{Date: 30/07/2021}  % For date in header

% REMOVE the % sign in the above lines if you want to
% compile this experiment file separately.

\begin{document}
	
	\chapter{Gamma Ray Spectrometer} % Replace the text with the title of the experiment
	\vspace{-1cm}
	\dateofexp{Date of Experiment: 30/07/2021} 
	% To display date in contents page
	
	\begin{center}% To display date of experiment in the title
		Date of Experiment: 30/07/2021
	\end{center}
	
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	% THE EXPERIMENT STARTS HERE %
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	
	\section{Aim}
	To calibrate a Gamma Ray Spectrometer.
	
	\section{Requirements}
	\begin{itemize}
		\item	Gamma source Sodium $ ^{22} $Na
		\item 	Gamma ray spectrometer
	\end{itemize}
	
	\section{Theory}
	In this experiment, we will study the radioactive decay of a nucleus by detecting gamma rays emitted consequent to the decay. Gamma ray detection is a multi-step process. It is possible to use GM counters for the detection of gamma radiation and for this purpose, the windowless tube is used. 
	
	Efficiency is just $ 1\% $ because of low gamma interaction of photon radiation with the material in the tube wall normally 1-2 mm of chrome steel, for producing electrons that can enter and ionize the gas. This is essential as there is a low gas density in the tube, and most high energy gamma photons pass through without being detected. 
		
\section{Procedure}
	
	\begin{enumerate}
		\item 	The given source is kept above the detector.
		\item 	The counter is set for 20 seconds, and voltage at zero counts are noted.
		
		\item 	Now vary the voltage by 50 and take the counts till \SI{350}{\volt}. Take two trials for each voltage.
		
		\item 	After \SI{350}{\volt}, increase the voltage in steps of \SI{10}{\volt} till about \SI{800}{\volt} and note the counts. Take two trials for each case.
		
		\item 	A graph of counts against voltage is plotted and the gamma ray spectrometer is calibrated.
		
	\end{enumerate}		


\section{Observations \& Graph}

\section{Result}

Gamma ray spectrometer is calibrated.

\end{document}
